---
title: "PA1_submission"
author: "Anzisca"
date: "07 December 2017"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
```


Reading the data into R
```{R read data}

setwd("C:/Users/anzisca.vanrooyen/Desktop/RepData_PeerAssessment1")
DATA <-read.csv("activity.csv")

```
Transform format to required type 

change date type to date class 
```{R transform data}
DATA$date <- as.Date(as.character(DATA$date))
```
Exclude Nulls
```{R Omit Null}
NAomit_DATA <-  na.omit(DATA)
```
## 1. Total number of steps taken per day
```{R MeansStepsByDay }
MeansStepsByDay <- setNames(aggregate(NAomit_DATA$steps, by=list(NAomit_DATA$date), sum),c("Date","Total_Steps"))
```
Histogram: Total number of steps taken per day
Load library

```{R load GGplot} 
library(ggplot2)
```

Histrogram plot

```{r Hist_MeansStepsByDay, echo=TRUE}
ggplot(MeansStepsByDay, aes(x = Total_Steps)) +
geom_histogram(fill = "yellow", binwidth=1000) +
labs(title = "Total Daily Steps", x = "Steps", y = "Frequency")
```

# 2. Mean and Median Total number of steps taken per day
Mean number of steps 

```{R MeanTotal_steps} 
mean(MeansStepsByDay$Total_Steps)
```
Median number of steps 

```{R Median_Total_steps} 
median(MeansStepsByDay$Total_Steps)
```

# 3. Average daily activity pattern  
```{r MeansStepsByInterval, echo=TRUE}
MeansStepsByInterval <- setNames(aggregate(NAomit_DATA$steps, by=list(NAomit_DATA$interval), sum),c("Interval","Total_Steps"))


ggplot(MeansStepsByInterval, aes(x = Interval, y=Total_Steps)) +
  labs(title = "Sum of Steps by Interval", x = "Interval", y = "Total_Steps")+
  geom_line(color="red") 

```
The time interval has the maximum number of steps: 

```{R MeanInt_Totsteps} 
MeansStepsByInterval[which.max(MeansStepsByInterval$Total_Steps),]
```

#4. Total number of steps , missing values substituted

Impute missing values 
Number of missing rows:
```{R Numberofmissings} 
nrow(is.na(DATA))
```

-Replace all missing values with aggregate measures
-Using na.aggregate from Zoo I am able to loop through the columns of the dataset
-Mean number of steps taken per data, Missing values substituted
```{R SubMissings}
library(zoo)
DATA2 <- DATA
DATA2[] <- lapply(DATA2, na.aggregate)

MeansStepsByDayInclNa <- setNames(aggregate(DATA2$steps, by=list(DATA2$date), sum),c("Date","Total_Steps"))
```
Histogram number of steps taken each day , Missing values substituted
Replacing the missing values with averages does not seem to change the average number of daily steps each day. Its does however moved the median towards the mean values
```{r MeansStepsByDayInclNa, echo=TRUE}
ggplot(MeansStepsByDayInclNa, aes(x = Total_Steps)) +
geom_histogram(fill = "Orange", binwidth=1000) +
labs(title = "Total Daily Steps", x = "Steps", y = "Frequency")
```
#5. Mean and Median Total number of steps taken per day , missing values substituted

Mean number of steps, Missing values substituted

```{R Mean_totInlcNA} 
mean(MeansStepsByDayInclNa$Total_Steps)
```
Median number of steps, Missing values substituted 

```{R Med_totInlcNA} 
median(MeansStepsByDayInclNa$Total_Steps)
```

#6. number of steps taken, averaged across all weekdays and weekends, missing values substituted
Add weekday and weekend functions 
```{R WeekdayCalc}
DATA2$weekday <- weekdays(DATA2$date)
DATA2$weekend <- ifelse (DATA2$weekday == "Saturday" | DATA2$weekday == "Sunday", "Weekend", "Weekday")
DATA2$weekpart <- as.factor(DATA2$weekend)
```

Panel plot  5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days
```{R MeanDataWeekendWeekday}
MeanDataWeekendWeekday <- aggregate(DATA2$steps, by=list(DATA2$weekpart, DATA2$interval), mean)
names(MeanDataWeekendWeekday)[1] ="weekpart"
names(MeanDataWeekendWeekday)[2] ="interval"
names(MeanDataWeekendWeekday)[3] ="steps"
```
```{r ggPlotMeanDataWeekendWeekday, echo=TRUE}
ggplot(MeanDataWeekendWeekday, aes(x = interval, y=steps, color=weekpart)) +
  geom_line() +
  facet_grid(weekpart ~ .) +
  labs(title = "Mean of Steps by Interval", x = "interval", y = "steps")
```

